Weather covariance matrices as U_hyp for GR50

### 2021-01-19 R/Analysis_v0.6_weather_covariance_Uhyp_greenup.Rmd
Edit manuscript to add cov matrix examples to Figure 2

ggcov
```

@@

-2*(m_midw_gr_no$loglik - m_midw_gr_hyp$loglik) # positive
-2*(m_midw_fl_no$loglik - m_midw_fl_hyp$loglik) # v. negative 
1 - pchisq(-2*(m_midw_fl_no$loglik - m_midw_fl_hyp$loglik), 30*32)
1 - pchisq(-2*(m_midw_gr_no$loglik - m_midw_gr_hyp$loglik), 15*32) # sig

-2*(m_gam_gr_no$loglik - m_gam_gr_hyp$loglik)  # positive
-2*(m_gam_fl_no$loglik - m_gam_fl_hyp$loglik) # v. positive
1 - pchisq(-2*(m_gam_fl_no$loglik - m_gam_fl_hyp$loglik), 30*32) # sig
1 - pchisq(-2*(m_gam_gr_no$loglik - m_gam_gr_hyp$loglik), 15*32) # sig

1-pchisq(0.0279,1) # LR -2(L0 - L1) gives LR of 0.0279 which gives p of 0.87
1-pchisq((6.89), 1) # LR of 6.89 gives p of ~0.01
```

@@

@@

f2b <- df_2gr_b %>%
  filter(Mass > 1e-05) %>%
  ggplot(aes(x = subpop, y = Mass)) + 
  geom_bar(aes(fill = Type), stat = "identity") +
  scale_fill_viridis_d(end = 0.9) +
  theme(axis.title.x=element_blank(),
        legend.position = "none",
        axis.text.x=element_text(hjust = 1, vjust = 1, angle = 45))
f2d <- df_2fl_b %>%
  filter(Mass > 1e-05) %>%
  ggplot(aes(x = subpop, y = Mass)) + 
  geom_bar(aes(fill = Type), stat = "identity") +
  scale_fill_viridis_d(end = 0.9) +
  theme(axis.title.x=element_blank(),
        legend.position = "none",
        axis.text.x=element_text(hjust = 1, vjust = 1, angle = 45))
f2a <- df_2gr_a %>%
  filter(Mass > 1e-06) %>%
  ggplot(aes(x = `Covariance Matrix`, y = Mass)) + 
  geom_bar(aes(fill = Type), stat = "identity") +
  facet_wrap(~subpop) +
  coord_flip() + scale_fill_viridis_d(end = 0.9) +
  theme(legend.position = c(-0.30, 0.3),
        axis.text.x=element_text(hjust = 1, vjust = 1, angle = 45), 
        panel.spacing.x = unit(-0, 'cm'))
f2c <- df_2fl_a %>%
  filter(Mass > 1e-06) %>%
  ggplot(aes(x = `Covariance Matrix`, y = Mass)) + 
  geom_bar(aes(fill = Type), stat = "identity") +
  facet_wrap(~subpop) +
  coord_flip() + scale_fill_viridis_d(end = 0.9) +
  theme(legend.position = "none",
        axis.text.x=element_text(hjust = 1, vjust = 1, angle = 45), 
        panel.spacing.x = unit(-0, 'cm'))

@@

>>>>

# ggcov
ggcov
```

@@

-2*(m_midw_gr_no$loglik - m_midw_gr_hyp$loglik) # positive
-2*(m_midw_fl_no$loglik - m_midw_fl_hyp$loglik) # v. negative 
1 - pchisq(-2*(m_midw_fl_no$loglik - m_midw_fl_hyp$loglik), 30*36)
1 - pchisq(-2*(m_midw_gr_no$loglik - m_midw_gr_hyp$loglik), 15*36) # sig

-2*(m_gam_gr_no$loglik - m_gam_gr_hyp$loglik)  # positive
-2*(m_gam_fl_no$loglik - m_gam_fl_hyp$loglik) # v. positive
1 - pchisq(-2*(m_gam_fl_no$loglik - m_gam_fl_hyp$loglik), 30*36) # sig
1 - pchisq(-2*(m_gam_gr_no$loglik - m_gam_gr_hyp$loglik), 15*36) # sig

1-pchisq(0.0279,1) # LR -2(L0 - L1) gives LR of 0.0279 which gives p of 0.87
1-pchisq((6.89), 1) # LR of 6.89 gives p of ~0.01
8+7+6+5+4+3+2+1
```

@@

```{r}
get_U_by_mass(m_gulf_gr_hyp,thresh = 0.01)
cov_gr1 <- mash_plot_Ulist(m_gulf_gr_hyp, range = 18, limits = TRUE)
cov_gr2 <- mash_plot_Ulist(m_gam_gr_hyp, 13, limits = TRUE)

fig_eqeff <- cov_gr1$equal_effects_df %>%
  ggplot(aes(x = .data$rowU, y = .data$colU)) +
  switchgrassGWAS::theme_oeco +
  geom_tile(aes(fill = .data$covar), na.rm = TRUE) +
  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                "white", "#27AD81FF", "#5DC863FF",
                                "#FDE725FF"),
                       limits = c(-1, 1)) +
  theme(legend.position = "none",
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.line.x = element_blank(),
        axis.line.y = element_blank(),
        axis.title.y.right = element_text()) +
  labs(x = "", y = "", subtitle = "equal_effects")

fig_ok_eff <- cov_gr2$Stand_BhatGulf_and_Midwest_GR50_OK_df %>%
  ggplot(aes(x = .data$rowU, y = .data$colU)) +
  switchgrassGWAS::theme_oeco +
  geom_tile(aes(fill = .data$covar), na.rm = TRUE) +
  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                "white", "#27AD81FF", "#5DC863FF",
                                "#FDE725FF"),
                       limits = c(-1, 1)) +
  theme(legend.position = "none",
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.line.x = element_blank(),
        axis.line.y = element_blank()) +
  xlab("") + ylab("") + ggtitle("single_effect_OK")
```

```{r second try making covar examples}
phe_hyp <- c("tave_10d", "tave_18d", "dyln_fl50", "cgdd_12c_gr2fl")

    if (i %in% c(1,2)) {
      cov_df <- phe_gr_df %>%
        ungroup() %>%
        left_join(select(sites, SITE, manu_site), by = "SITE") %>%
        filter(SUBPOP %in% subpop_v2[[k]]) %>%
        select(PLANT_ID, SUBPOP, manu_site, phe_hyp[i]) %>%
        group_by(PLANT_ID, SUBPOP, manu_site) %>%
        mutate(IND = row_number()) %>%
        pivot_wider(names_from = manu_site, values_from = phe_hyp[i]) %>%
        mutate(PLANT_ID = as.factor(PLANT_ID),
               IND = as.factor(IND)) %>%
        select(-IL) %>%
        select(PLANT_ID, SUBPOP, IND, TX1, TX2, TX3, OK, MO, NE, MI, SD)
    } else {
      cov_df <- phe_fl_df %>%
        ungroup() %>%
        left_join(select(sites, SITE, manu_site), by = "SITE") %>%
        filter(SUBPOP %in% subpop_v2[[k]]) %>%
        select(PLANT_ID, SUBPOP, manu_site, phe_hyp[i]) %>%
        group_by(PLANT_ID, SUBPOP, manu_site) %>%
        mutate(IND = row_number()) %>%
        pivot_wider(names_from = manu_site, values_from = phe_hyp[i]) %>%
        mutate(PLANT_ID = as.factor(PLANT_ID),
               IND = as.factor(IND)) %>%
        select(-IL) %>%
        select(PLANT_ID, SUBPOP, IND, TX1, TX2, TX3, OK, MO, NE, MI, SD)
    }

  cor_phe <- cor(cov_df[,-(1:3)], use = "pairwise")
  diag(cor_phe) <- cov_df %>%
    ungroup() %>%
    summarise(TX1 = sqrt(var(TX1, na.rm = TRUE))/mean(TX1, na.rm = TRUE),
              TX2 = sqrt(var(TX2, na.rm = TRUE))/mean(TX2, na.rm = TRUE),
              TX3 = sqrt(var(TX3, na.rm = TRUE))/mean(TX3, na.rm = TRUE),
              OK = sqrt(var(OK, na.rm = TRUE))/mean(OK, na.rm = TRUE),
              MO = sqrt(var(MO, na.rm = TRUE))/mean(MO, na.rm = TRUE),
              NE = sqrt(var(NE, na.rm = TRUE))/mean(NE, na.rm = TRUE),
              MI = sqrt(var(MI, na.rm = TRUE))/mean(MI, na.rm = TRUE),
              SD = sqrt(var(SD, na.rm = TRUE))/mean(SD, na.rm = TRUE)) %>%
    as_vector()

  if(isSymmetric(cor_phe)){
        for(m in 1:nrow(cor_phe)){
          for(j in 1:ncol(cor_phe)){
            if(m < j){
              cor_phe[m, j] <- NA
            }
          }
        }
    }
  U1 <- as_tibble(cor_phe, rownames = "rowU", .name_repair = "unique") %>%
          pivot_longer(cols = -.data$rowU, names_to = "colU",
                       values_to = "covar") %>%
          filter(!is.na(.data$covar))
  U1$colU <- factor(U1$colU, levels = colnames(cor_phe))
  U1$rowU <- factor(U1$rowU, levels = colnames(cor_phe))
  U1 <- U1 %>%
    mutate(phe = phe_hyp[i],
           subpop = names(subpop_v2)[k],
           subpop = case_when(subpop == "Gulf_and_Midwest" ~ "Both",
                              TRUE ~ subpop))

  if(i == 1 & k == 1){
    cov_flex <- U1
  } else {
    cov_flex <- cov_flex %>%
      full_join(U1)
  }
    }
}

cov_flex$phe <- factor(cov_flex$phe, levels = c("tave_18d", "tave_10d", "dyln_fl50", "cgdd_12c_gr2fl"))

figure2_a <- cov_flex %>%
  filter(phe %in% c("tave_10d", "tave_18d")) %>%
  ggplot(aes(x = .data$rowU, y = .data$colU)) +
  switchgrassGWAS::theme_oeco +
  geom_tile(aes(fill = .data$covar), na.rm = TRUE) +
  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                "white", "#27AD81FF", "#5DC863FF",
                                "#FDE725FF"),
                       limits = c(-1, 1)) +
  #geom_text(aes(label = round(.data$covar, 1)), color = "darkgrey") +
  facet_grid(rows = vars(phe), cols = vars(subpop)) + 
  theme(legend.position = "right",
        #axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.line.x = element_blank(),
        axis.line.y = element_blank(),
        panel.spacing.x = unit(-0, 'cm')) +
  xlab("") + ylab("")

legend <- get_legend(figure2_a)  
figure2_a <- cov_flex %>%
  filter(phe %in% c("tave_10d", "tave_18d")) %>%
  ggplot(aes(x = .data$rowU, y = .data$colU)) +
  switchgrassGWAS::theme_oeco +
  geom_tile(aes(fill = .data$covar), na.rm = TRUE) +
  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                "white", "#27AD81FF", "#5DC863FF",
                                "#FDE725FF"),
                       limits = c(-1, 1)) +
  #geom_text(aes(label = round(.data$covar, 1)), color = "darkgrey") +
  facet_grid(rows = vars(phe), cols = vars(subpop)) + 
  theme(legend.position = "none",
        #axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.line.x = element_blank(),
        axis.line.y = element_blank(),
        panel.spacing.x = unit(-0, 'cm')) +
  xlab("") + ylab("")

figure2_d <- cov_flex %>%
  filter(phe %in% c("cgdd_12c_gr2fl", "dyln_fl50")) %>%
  ggplot(aes(x = .data$rowU, y = .data$colU)) +
  switchgrassGWAS::theme_oeco +
  geom_tile(aes(fill = .data$covar), na.rm = TRUE) +
  scale_fill_gradientn(colors = c("#440154FF", "#3B528BFF", "#2C728EFF",
                                "white", "#27AD81FF", "#5DC863FF",
                                "#FDE725FF"),
                       limits = c(-1, 1)) +
  #geom_text(aes(label = round(.data$covar, 1)), color = "darkgrey") +
  facet_grid(rows = vars(phe), cols = vars(subpop)) + 
  theme(legend.position = "none",
        #axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.line.x = element_blank(),
        axis.line.y = element_blank(),
        panel.spacing.x = unit(-0, 'cm')) +
  xlab("") + ylab("")
```

@@

f2b <- df_2gr_b %>%
  filter(Mass > 1e-05) %>%
  ggplot(aes(x = subpop, y = Mass)) + 
  geom_bar(aes(fill = Type), stat = "identity") +
  scale_fill_viridis_d(end = 0.9) +
  theme(axis.title.x=element_blank(),
        legend.position = "none",
        axis.text.x=element_text(hjust = 1, vjust = 1, angle = 45))
f2d <- df_2fl_b %>%
  filter(Mass > 1e-05) %>%
  ggplot(aes(x = subpop, y = Mass)) + 
  geom_bar(aes(fill = Type), stat = "identity") +
  scale_fill_viridis_d(end = 0.9) +
  theme(axis.title.x=element_blank(),
        legend.position = "none",
        axis.text.x=element_text(hjust = 1, vjust = 1, angle = 45))
f2a <- df_2gr_a %>%
  filter(Mass > 1e-06) %>%
  ggplot(aes(x = `Covariance Matrix`, y = Mass)) + 
  geom_bar(aes(fill = Type), stat = "identity") +
  facet_wrap(~subpop) +
  coord_flip() + scale_fill_viridis_d(end = 0.9) +
  theme(legend.position = c(-0.4, 0.3),
        axis.text.x=element_text(hjust = 1, vjust = 1, angle = 45), 
        panel.spacing.x = unit(-0, 'cm')) #+ labs(x = "")
f2c <- df_2fl_a %>%
  filter(Mass > 1e-06) %>%
  ggplot(aes(x = `Covariance Matrix`, y = Mass)) + 
  geom_bar(aes(fill = Type), stat = "identity") +
  facet_wrap(~subpop) +
  coord_flip() + scale_fill_viridis_d(end = 0.9) +
  theme(legend.position = "none",
        axis.text.x=element_text(hjust = 1, vjust = 1, angle = 45), 
        panel.spacing.x = unit(-0, 'cm')) #+ labs(x = "")

@@

```{r}
ggdraw() +
  draw_plot(figure2_a, x = -.02, y = 0.51, width = 0.32, height = 0.35) +
  draw_plot(fig_eqeff, x = 0.02, y = 0.35, width = 0.125, height = 0.2) +
  draw_plot(figure2_d, x = -.02, y = 0.03, width = 0.32, height = 0.35) +
  draw_plot(f2a, x = 0.3, y = 0.47, width = 0.55, height = 0.53) +
  draw_plot(legend, x = 0.15, y = 0.34, width = 0.1, height = 0.2) +
  draw_plot(f2b, x = 0.85, y = 0.47, width = 0.15, height = 0.53) +
  draw_plot(f2c, x = 0.3, y = 0, width = 0.55, height = 0.47) +
  draw_plot(f2d, x = 0.85, y = 0, width = 0.15, height = 0.47)

save_plot(filename = "Figure_2_Covariance_matrices_and_posterior_weights.png", plot = last_plot(), base_height = 8, base_width = 9)
```
